### 2022-01-28 supp/03-gm.analysis.Rmd
<edit gm>

# Geometric morphometrics

## Load packages + data

```{r load.data, echo = TRUE}
# load packages
#devtools::install_github("mlcollyer/RRPP")
#devtools::install_github("geomorphR/geomorph", ref = "Stable", build_vignettes = TRUE)
library(here)
library(geomorph)
library(tidyverse)
library(wesanderson)

# read GM data
source('readmulti.csv.R')
setwd("./data")
filelist <- list.files(pattern = ".csv")
coords <- readmulti.csv(filelist)
setwd("../")

# read qualitative data
qdata <- read.csv("qdata.csv", 
                  header = TRUE, 
                  row.names = 1)

qdata <- qdata[match(dimnames(coords)[[3]],
                     rownames(qdata)),]
```

## Generalised Procrustes Analysis

Landmark data were aligned to a global coordinate system [@RN11622;@RN11623;@RN11563], achieved through generalised Procrustes superimposition [@RN478] performed in R 4.0.3 [@R] using the `geomorph` library v. 3.3.2 [@RN11530;@RN1774]. Procrustes superimposition translates, scales, and rotates the coordinate data to allow for comparisons among objects [@RN11564;@RN478]. The `geomorph` package uses a partial Procrustes superimposition that projects the aligned specimens into tangent space subsequent to alignment in preparation for the use of multivariate methods that assume linear space [@RN1646;@RN11563].

```{r gpa, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# gpa
Y.gpa <- gpagen(coords, 
                PrinAxes = TRUE, 
                ProcD = TRUE, 
                Proj = TRUE, 
                print.progress = FALSE)

# geomorph data frame
gdf <- geomorph.data.frame(shape = Y.gpa$coords, 
                           size = Y.gpa$Csize,
                           type = qdata$type)

# render 3d gpa plot
plot(Y.gpa)

# gpa plot
#knitr::include_graphics('images/gpa3d.png')

# add centroid size to qdata
qdata$csz <- Y.gpa$Csize

# print updated qdata with centroid size
knitr::kable(qdata, 
             align = "lccccc", 
             caption = "Modified attributes included in qdata.")
```

### Boxplot

```{r box1, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap = "Boxplot of centroid size by spatial/temporal unit."}
# attributes for boxplot
csz <- qdata$csz
group <-  qdata$group
type <- qdata$type

# boxplot of vessel centroid size by type
csz.type <- ggplot(qdata, aes(x = type, y = csz, color = type)) + 
  geom_boxplot(notch = FALSE) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.3) +
  scale_colour_manual(values = wes_palette("Moonrise2")) +
  theme(legend.position = "none") +
  labs(x = 'Group', y = 'Centroid Size')

# render plot
csz.type
```

## Principal Components Analysis

Principal components analysis [@RN1746] was used to visualise shape variation among the bottles. The shape changes described by each principal axis are commonly visualised using thin-plate spline warping of a reference 3D mesh [@RN1731;@RN479].

```{r gmpca, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# principal components analysis
pca<-gm.prcomp(Y.gpa$coords)
summary(pca)

# set plot parameters (types)
pch.gps.type <- c(15,17)[as.factor(type)]
col.gps.type <- wes_palette("Moonrise2")[as.factor(type)]
col.hull.type <- c("#798E87","#C27D38")

# plot pca by comb
pc.plot.type <- plot(pca, 
                 asp = 1,
                 pch = pch.gps.type,
                 col = col.gps.type)
                    shapeHulls(pc.plot.type, 
                             groups = type,
                             group.cols = col.hull.type)
```

```{r pca.warp, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Results of PCA summarising shape variation in the aggregated sample; gray, north-formative/early; orange, north-late/historic; brown, south-formative/early; and black, south-late/historic; with mean shapes (gray) contrasted with max/min shapes (black)."}
# pca warp                   
#knitr::include_graphics('images/pca-warp.jpg')
```

## Vessel allometry

```{r allometry.1, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# allometry: shape as a function of vessel size?
fit.size<-procD.lm(shape ~ size, 
                   data = gdf, 
                   print.progress = FALSE, 
                   iter = 9999)

# Procrustes ANOVA: does shape change with size?
anova(fit.size)
```

```{r common.allometry.1, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# common allometries
fit.common<-procD.lm(shape ~ size + type, 
                     data = gdf, 
                     print.progress = FALSE, 
                     iter = 9999)

anova(fit.common)

common.allometry <- plot(fit.size,
                         type = "regression",
                         reg.type = "RegScore",
                         predictor = log(gdf$size),
                         pch = pch.gps.type,
                         col = col.gps.type)

# common allometric component (Mitteroecker 2004)
plotAllometry(fit.size, 
              size = gdf$size, 
              logsz = TRUE, 
              method = "CAC", 
              pch = pch.gps.type, 
              col = col.gps.type)
```

```{r unique.allometry.1, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# unique allometries
fit.unique<-procD.lm(shape ~ size * type, 
                     data = gdf, 
                     print.progress = FALSE, 
                     iter = 9999)

anova(fit.unique)

# do vessel forms express parallel, convergent, or divergent morphological characteristics?
plotAllometry(fit.unique, 
              size = gdf$size, 
              logsz = TRUE, 
              method = "PredLine", 
              pch = pch.gps.type, 
              col = col.gps.type)
```

## Vessel size and shape

A residual randomisation permutation procedure (RRPP; n = 10,000 permutations) was used for all Procrustes ANOVAs [@RN1655;@RN11775], which has higher statistical power and a greater ability to identify patterns in the data should they be present [@RN1719]. To assess whether shape changes differ by group (geography and time), Procrustes ANOVAs [@RN1749] were also run that enlist effect-sizes (zscores) computed as standard deviates of the generated sampling distributions [@RN1756].

```{r def.mod.1, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# size as a function of vessel form (bottle/jar)?
fit.size.type <- procD.lm(size ~ type,
                data = gdf,
                print.progress = FALSE,
                iter = 9999)

# Procrustes ANOVA: do sizes differ as a function of vessel form?
anova(fit.size.type)

# shape as a function of vessel form?
fit.shape.type <- procD.lm(shape ~ type,
                data = gdf,
                print.progress = FALSE,
                iter = 9999)

# Procrustes ANOVA: do shapes differ as a function of vessel form?
anova(fit.shape.type)
```

## Morphological disparity

```{r disparity, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# morphological disparity: do either of the groups display greater 
# size variation among individuals relative to the other group?
morphol.disparity(fit.size.type, 
                  groups = qdata$type, 
                  data = gdf, 
                  print.progress = FALSE, 
                  iter = 9999)

# morphological disparity: do either of the groups display greater 
#shape variation among individuals relative to the other group?
morphol.disparity(fit.shape.type, 
                  groups = qdata$type, 
                  data = gdf, 
                  print.progress = FALSE, 
                  iter = 9999)
```

## Mean shapes

```{r mshape.h1, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# subset landmark coordinates to produce mean shapes
new.coords<-coords.subset(A = Y.gpa$coords,
                          group = qdata$type)
names(new.coords)

# group shape means
mean <- lapply(new.coords, mshape)

# plot mean shapes (type)
plot(mean$bottle)
plot(mean$jar)

# comparison plots
plotRefToTarget(mean$bottle, 
                mean$jar, 
                method = c("points"),
                mag = 1)

knitr::include_graphics('images/mshape.png')
```