### 2022-01-28 supp/04-gm.sculptural.Rmd
<init .s>

# Sculptural elements

## Load packages + data

```{r load.data.s, echo = TRUE}
# load packages
#devtools::install_github("mlcollyer/RRPP")
#devtools::install_github("geomorphR/geomorph", ref = "Stable", build_vignettes = TRUE)
library(here)
library(geomorph)
library(tidyverse)
library(wesanderson)

# read GM data
source('readmulti.csv.R')
setwd("./data")
filelist <- list.files(pattern = ".csv")
coords <- readmulti.csv(filelist)
setwd("../")

# read qualitative data
qdata <- read.csv("qdata.csv", 
                  header = TRUE, 
                  row.names = 1)

qdata <- qdata[match(dimnames(coords)[[3]],
                     rownames(qdata)),]

qdata <- qdata %>% 
  unite("merged", group:type, remove = FALSE)
```

## Generalised Procrustes Analysis

Landmark data were aligned to a global coordinate system [@RN11622;@RN11623;@RN11563], achieved through generalised Procrustes superimposition [@RN478] performed in R 4.0.3 [@R] using the `geomorph` library v. 3.3.2 [@RN11530;@RN1774]. Procrustes superimposition translates, scales, and rotates the coordinate data to allow for comparisons among objects [@RN11564;@RN478]. The `geomorph` package uses a partial Procrustes superimposition that projects the aligned specimens into tangent space subsequent to alignment in preparation for the use of multivariate methods that assume linear space [@RN1646;@RN11563].

```{r gpa.s, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# gpa
Y.gpa <- gpagen(coords, 
                PrinAxes = TRUE, 
                ProcD = TRUE, 
                Proj = TRUE, 
                print.progress = FALSE)

# geomorph data frame
gdf <- geomorph.data.frame(shape = Y.gpa$coords, 
                           size = Y.gpa$Csize,
                           merged = qdata$merged)

# render 3d gpa plot
plot(Y.gpa)

# gpa plot
#knitr::include_graphics('images/gpa3d.png')

# add centroid size to qdata
qdata$csz <- Y.gpa$Csize

# print updated qdata with centroid size
knitr::kable(qdata, 
             align = "lccccc", 
             caption = "Modified attributes included in qdata.")
```

### Boxplot

```{r box1.s, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap = "Boxplot of centroid size by spatial/temporal unit."}
# attributes for boxplot
csz <- qdata$csz
merged <-  qdata$merged

# palette
pal <- wes_palette("Moonrise2", 6, type = "continuous")

# boxplot of vessel centroid size by merged
csz.merged <- ggplot(qdata, aes(x = merged, y = csz, color = merged)) + 
  geom_boxplot(notch = FALSE) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.3) +
  scale_colour_manual(values = pal) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x = 'Sculptural Element + Vessel Form', 
       y = 'Centroid Size')

# render plot
csz.merged
```

## Principal Components Analysis

Principal components analysis [@RN1746] was used to visualise shape variation among the bottles. The shape changes described by each principal axis are commonly visualised using thin-plate spline warping of a reference 3D mesh [@RN1731;@RN479].

```{r gmpca.s, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# principal components analysis
pca<-gm.prcomp(Y.gpa$coords)
summary(pca)

# set plot parameters (mergeds)
pch.gps.merged <- c(1:6)[as.factor(merged)]

# plot pca by merged
plot(pca,
     asp = 1,
     pch = pch.gps.merged,
     col = pal)
```

```{r pca.warp.s, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE, fig.cap="Results of PCA summarising shape variation in the aggregated sample; gray, north-formative/early; orange, north-late/historic; brown, south-formative/early; and black, south-late/historic; with mean shapes (gray) contrasted with max/min shapes (black)."}
# pca warp                   
#knitr::include_graphics('images/pca-warp.jpg')
```

## Vessel size and shape

A residual randomisation permutation procedure (RRPP; n = 10,000 permutations) was used for all Procrustes ANOVAs [@RN1655;@RN11775], which has higher statistical power and a greater ability to identify patterns in the data should they be present [@RN1719]. To assess whether shape changes differ by merged (geography and time), Procrustes ANOVAs [@RN1749] were also run that enlist effect-sizes (zscores) computed as standard deviates of the generated sampling distributions [@RN1756].

```{r def.mod.1.s, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# size as a function of vessel form (bottle/jar)?
fit.size.merged <- procD.lm(size ~ merged,
                            data = gdf,
                            print.progress = FALSE,
                            iter = 9999)

# Procrustes ANOVA: do sizes differ as a function of vessel form?
anova(fit.size.merged)

# pairwise comparison of LS means = which differ?
pairwise.size <- pairwise(fit.size.merged,
                          groups = qdata$merged)
summary(pairwise.size, 
        confidence = 0.95, 
        test.type = "dist")

# shape as a function of vessel form?
fit.shape.merged <- procD.lm(shape ~ merged,
                data = gdf,
                print.progress = FALSE,
                iter = 9999)

# Procrustes ANOVA: do shapes differ as a function of vessel form?
anova(fit.shape.merged)

# pairwise comparison of LS means = which differ?
pairwise.shape <- pairwise(fit.shape.merged,
                           groups = qdata$merged)
summary(pairwise.shape, 
        confidence = 0.95, 
        test.type = "dist")
```

## Morphological disparity

```{r disparity.s, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# morphological disparity: do either of the mergeds display greater 
# size variation among individuals relative to the other merged?
morphol.disparity(fit.size.merged, 
                  groups = qdata$merged, 
                  data = gdf, 
                  print.progress = FALSE, 
                  iter = 9999)

# morphological disparity: do either of the mergeds display greater 
#shape variation among individuals relative to the other merged?
morphol.disparity(fit.shape.merged, 
                  groups = qdata$merged, 
                  data = gdf, 
                  print.progress = FALSE, 
                  iter = 9999)
```

## Mean shapes

```{r mshape.s, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# subset landmark coordinates to produce mean shapes
new.coords<-coords.subset(A = Y.gpa$coords,
                          group = qdata$merged)
names(new.coords)

# merged shape means
mean <- lapply(new.coords, mshape)

# plot mean shapes (merged)
plot(mean$anthropomorph_bottle)
plot(mean$anthropomorph_jar)

plot(mean$phytomorph_bottle)
plot(mean$phytomorph_jar)

plot(mean$zoomorph_bottle)
plot(mean$zoomorph_jar)

# comparison plots
plotRefToTarget(mean$anthropomorph_bottle, 
                mean$zoomorph_jar, 
                method = c("points"),
                mag = 1)

#knitr::include_graphics('images/mshape.png')
```